---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# prorata

<!-- badges: start -->

<!-- badges: end -->

The goal of prorata is to ...

## Installation

You can install the development version of prorata from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("UchidaMizuki/prorata")
```

## Example

```{r, message=FALSE, warning=FALSE}
library(prorata)
library(tidyverse)
```

### Freeny's Revenue Data

-   Quarterly revenues are estimated from prorated values for annual revenues based on price index, income level, and market potential
-   Estimate weights that sum to 1 with the least squared error (or absolute error)
-   Input the following `y` and `x` to `prorata()`
    -   `y`: Quarterly revenues (observations x categories)
    -   `x`: Prorated values for annual revenues (observations x categories x clusters)
-   Apply `predict()` to `prorata()` output

```{r}
data <- freeny |> 
  as_tibble(rownames = "year_quarter") |> 
  separate_wider_delim(year_quarter, 
                       delim = ".",
                       names = c("year", "quarter"),
                       too_few = "align_start") |> 
  replace_na(list(quarter = "0"))

y <- xtabs(y ~ year + quarter,
           data = data)
y

x <- data |> 
  mutate(across(!c(quarter, y),
                \(x) x / sum(x) * sum(y)),
         .by = year,
         .keep = "unused") |> 
  pivot_longer(!c(year, quarter)) |> 
  xtabs(value ~ year + quarter + name,
        data = _)
x

fitted <- prorata(y, x,
                  type = "squared")
fitted

pred <- predict(fitted,
                new_data = x)
pred

tibble(obs = as.double(y),
       pred = as.double(pred)) |>
  filter(obs > 0) |>
  ggplot(aes(obs, pred)) +
  geom_point()
```
